---
title: "Homework 7"
author: "Jacob Carey"
date: \today
header-includes:
    - \usepackage{amsthm}
    - \usepackage{mathtools}
output: pdf_document
---

```{r libs, echo=FALSE}
library(ggplot2)
```

1.

2.
    (i)
        The second derivative of $e^x$ is $e^x$, which is positive on the interval $(-\infty, \infty)$ and is thus convex.
        ```{r p2-i}
        x <- seq(-10, 10, by=0.1)
        qplot(x, exp(x), geom="line")
        ```

    (ii)
        The second derivative of $x^2$ is $2$, which is positive on the interval $(-\infty, \infty)$ and is thus convex.
        ```{r p2-ii}
        qplot(x, x^2, geom="line")
        ```

    (iii)
        The second derivative of $\ln x$ is $-\frac{1}{x^2}$, which is not postive on the interval $(-\infty, \infty)$ and is thus not convex.
        ```{r p2-iii}
        x <- seq(0, 10, by=0.1)
        qplot(x, log(x), geom="line")
        ```

3.

4.
    ```{r p4}
    x <- seq(-0.99, 1.99, by=0.01)
    g_x <- (1 + x) * log(1 + x)
    R_x <- (x ^ 2) / (2 * (1 + x / 3))
    data <- rbind(data.frame(x=x, y=g_x, label="g(x)"),
                  data.frame(x=x, y=x+R_x, label="x+R(x)"))
    ggplot(data, aes(x=x, y=y)) +
        geom_line(aes(colour=label, linetype=label))
    ```
    This is a pretty good approximation

5. Given that each X_i is $p$ and the variance is $p(1-p)$, then the mean and variance of $\bar{X_n}$ is also $p$ and $p(1-p)$. Then by Chebyshev's inequality (and by extension, the weak law  of large numbers)
   $$
   P(|\bar{X_n}-p|<\epsilon) \geq 1 - \frac{p(1-p)}{n\epsilon^2}
   $$
   If we want the probability of the inequality to be at least 0.95, we solve for n
   $$
   \begin{aligned}
   1-\frac{p(1-p)}{n\times 0.05^2} = 0.95 \\
   n - 0.95n = \frac{p(1-p)}{0.05^2} \\
   \boxed{n=\frac{p(1-p)}{0.05^3}}
   \end{aligned}
   $$
   In election polling, this is useful for determining a sample size such that with frequency 0.95, the sample mean will be within $0.05$ of the true proportion.

